---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
act <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
```{r message = FALSE}
library(dplyr)
library(ggplot2)
stepsPerDay <-act %>% filter(!is.na(steps)) %>% group_by(date) %>% summarise(steps = sum(steps))
ggplot(data=stepsPerDay, aes(x=steps)) + geom_histogram() + ggtitle("Distribution of total number of steps taken per day")
```


```{r}
res <-stepsPerDay  %>% summarise(meanSteps = mean(steps), medianSteps = median(steps) )
```

The mean total number of steps taken per day is `r format(round(res$meanSteps,2), nsmall=2)` and the median is `r format(round(res$medianSteps,2), nsmall=2)`.

## What is the average daily activity pattern?

```{r}
dat <- act %>% filter(!is.na(steps)) %>% group_by(interval) %>% summarise(meansteps = mean(steps))
plot(x=dat$interval, y=dat$meansteps, type = "l", xlab = "5-minute interval", ylab = "average number of steps taken", main = "Average daily activity pattern")
```

The maximum average number of steps is `r format(round(dat$meansteps[which.max(dat$meansteps)],2), nsmall = 2)` during interval `r dat$interval[which.max(dat$meansteps)]` 

## Imputing missing values

```{r}
nmissing <- sum(is.na(act$steps))
```

There are `r nmissing` missing values in the dataset

```{r message = FALSE}
# imputing the mean value for that 5-minute interval for missing values
# The mean value for each interval is already computed in the dat dataframe
actf <- act %>% mutate(steps = ifelse(is.na(steps), round(dat$meansteps[dat$interval == interval]), steps))
stepsPerDay2 <- actf %>% group_by(date) %>% summarise(steps = sum(steps))
ggplot(data=stepsPerDay2, aes(x=steps)) + geom_histogram() + ggtitle("Distribution of total number of steps taken per day after imputation")
```


```{r}
res2 <- stepsPerDay2  %>% summarise(meanSteps = mean(steps), medianSteps = median(steps) )
```

The mean total number of steps taken per day after imputation is `r format(round(res2$meanSteps,2), nsmall=2)` and the median is `r format(round(res$medianSteps2,2), nsmall=2)`.


## Are there differences in activity patterns between weekdays and weekends?
